---
title: "hw6"
author: "Sean Sorek"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      fig.align = "center",
                      fig.width = 8,
                      message = F,
                      warning = F)

library(tidyverse)
```

# Problem 2

```{r}
homicide <- read_csv("data/homicide-data.csv") |>
  mutate(
    city_state = stringr::str_c(city, state, sep = ", "),
    unsolved = map_lgl(
      disposition,
      \(i) str_detect(i, "Closed without arrest") ||
        str_detect(i, "Open/No arrest")
      ),
    victim_age = as.integer(victim_age)) |> 
  filter(!(city_state %in% c(
    "Dallas, TX",
    "Pheonix, AZ",
    "Kansas City, MO",
    "Tulsa, AL"
  )),
  victim_race %in% c("White", "Black")) |>
  na.omit()
  
homicide |>
  filter(city_state == "Baltimore, MD") |>
  glm(unsolved ~ victim_age + victim_sex + victim_race, data = _, family = "binomial") |> 
  broom::tidy() |>
  mutate(
    lower_coef = estimate - 1.96*std.error,
    upper_coef = estimate + 1.96*std.error,
    lower_odds = exp(lower_coef),
    upper_odds = exp(upper_coef),
    odds_ratio = exp(estimate)
    ) |> 
  filter(term == "victim_sexMale") |>
  select(lower_odds, odds_ratio, upper_odds)

get_gender_odds_ratio <- function(x){
  x |>
    glm(unsolved ~ victim_age + victim_sex + victim_race, data = _, family = "binomial") |> 
  broom::tidy() |>
  mutate(
    lower_coef = estimate - 1.96*std.error,
    upper_coef = estimate + 1.96*std.error,
    lower_odds = exp(lower_coef),
    upper_odds = exp(upper_coef),
    odds_ratio = exp(estimate)
    ) |> 
  filter(term == "victim_sexMale") |>
  select(lower_odds, odds_ratio, upper_odds)
}
homicide |>
  nest(data = -city_state) |>
  mutate(
    results = map(data, get_gender_odds_ratio)
  ) |>
  unnest(results) |> select(-data) |>
  mutate(
    city_state = as.factor(city_state),
    city_state = fct_reorder(city_state, odds_ratio, .desc = T)
  ) |>
  ggplot(aes(city_state, y = odds_ratio,
             ymin = lower_odds, ymax = upper_odds)) + geom_errorbar() + geom_point() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -75, hjust=0)) +
  labs(x = "City", 
       y = "Odds Ratio",
       title = "Effect of Male Gender on Odds of Crime being Unsolved by City")
  

```

# Problem 3
## Data Cleaning
```{r}
birthweight <- read_csv("data/birthweight.csv") |>
  mutate(
    babysex = factor(babysex, labels = c("male", "female")),
    frace = factor(frace, labels = c(
      "White",
      "Black",
      "Asian",
      "Puerto Rican",
      "Other"
    )),
    malform = factor(malform, labels = c(
      "absent", "present"
    )),
    mrace = factor(mrace, labels = c(
      "White",
      "Black",
      "Asian",
      "Puerto Rican"
    ))
  )


```

## Model Selection
```{r}
step_lm <- step(lm(bwt ~ ., birthweight), direction = "both", trace = 0)
```

I used step-wise selection with AIC since I have no prior beliefs as to what predictors would significantly affect birth weight.

```{r}
birthweight |> 
  modelr::add_residuals(step_lm) |>
  modelr::add_predictions(step_lm) |>
  ggplot(aes(pred, resid)) + geom_point() +
  theme_bw() +
  labs(x = "Predicted Value", 
       y = "Residual",
       title = "Residuals vs Predicted Value for Stepwise AIC Selected Model")
```

The model appears to be appropriate for birth weights above 2kgs. Since most of the data is in the appropriate range, this model is acceptable.

## Model Comparison
```{r}
birthweight |>
  modelr::crossv_mc(100) |>
  mutate(
    step_model = map(train, \(df) lm(formula(step_lm), data = df)),
    main_model = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    int_model = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))
  ) |>
  mutate(
    rmse_step = map2_dbl(
      step_model, test, \(mod, df) modelr::rmse(model = mod, data = df)),
    rmse_main = map2_dbl(
      main_model, test, \(mod, df) modelr::rmse(model = mod, data = df)),
    rmse_int = map2_dbl(
      int_model, test, \(mod, df)   modelr::rmse(model = mod, data = df))) |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  theme_bw() +
  labs(title = "RMSE of Different Models")
```

The model chosen using step-wise selection performs far better than either candidate model.